Fast and adaptive cointegration based model for forecasting financial time series

### 2017-12-09 slides/defense-slides.tex
primera contribucion lista

\documentclass{beamer}
%\usepackage[spanish,activeacute]{babel}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{color}
\usepackage{tikz}
\usepackage{array}
%\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{booktabs}
\usetikzlibrary{shapes,arrows}
%\usepackage{multicol}
\usepackage{hhline}
\usepackage{pgf}
\usepackage{algorithm} 
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\def\R{{\mathbb R }}
\def\N{{\mathbb N }}

\usetheme{metropolis}

\definecolor{BlueTOL}{HTML}{222255}
\definecolor{BrownTOL}{HTML}{666633}
\definecolor{GreenTOL}{HTML}{225522}
\setbeamercolor{normal text}{fg=BlueTOL,bg=white}
\setbeamercolor{alerted text}{fg=BrownTOL}
\setbeamercolor{example text}{fg=GreenTOL}

@@

\setbeamercolor{block title alerted}{use=alerted text,
    fg=alerted text.fg,
    bg=alerted text.bg!80!alerted text.fg}
\setbeamercolor{block body alerted}{use={block title alerted, alerted text},
    fg=alerted text.fg,
    bg=block title alerted.bg!50!alerted text.bg}
\setbeamercolor{block title example}{use=example text,
    fg=example text.fg,
    bg=example text.bg!80!example text.fg}
\setbeamercolor{block body example}{use={block title example, example text},
    fg=example text.fg,
    bg=block title example.bg!50!example text.bg}

@@

\author[Paola Arce]{Paola Arce \\ Departamento de Inform\'atica, UTFSM \\ Thesis advisor: Luis Salinas}
\date{December 11, 2017}
\title[Thesis defense]{Fast and adaptive cointegration based model for forecasting financial time series}
\subtitle{Seminario Inform\'atica CCTVal}

\titlegraphic{
    \tikz[overlay,remember picture]
        \node[at=(current page.south east), anchor=south east] {
            \includegraphics[height=1.5cm]{img/DIT}
        };
}

@@

\begin{frame}[plain]
\titlepage
\end{frame}
%\usebackgroundtemplate{ }

\begin{frame}
\frametitle{Outline}
\begin{itemize}
\item Motivation: Financial time series and HFT
\item Cointegration and Integration concepts
\item VECM model
\item Proposal
\item Experiments
\item Conclusions
\end{itemize}
\end{frame}

\section{Introduction}

@@

\begin{frame}
\frametitle{High Frequency data}
High frequency time series are commonly:
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Non-stationary
\item Irregularly spaced over time.
%\item Heterogeneity
\item Heteroscedastic
%\item Volatility
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=0.35\paperwidth]{img/forecast}
\end{figure}
\end{column}
\end{columns}
\vspace{5mm}
This imply that classical statistical models can't be used. Therefore, financial time series forecasting is still a challenge.
\end{frame}

@@

\begin{frame}
\frametitle{Stationarity: consequences}
One of the consequences of stationary processes is how they recover from a shock. A shock represents an unexpected change in a variable or a in its error term in a particular period of time. For stationary time series, shocks to the system will gradually die away. 
\end{frame}

\begin{frame}
\frametitle{Non-stationary processes}
There are different types of non-stationary time series models often found in economics:
\begin{block}{Stochastic trend}
Stochastic trend processes are also called {\color{red}unit root or I(1)} if they have the following form:
{\color{blue}
\[
y_t = \mu + y_{t-1} + \epsilon_t
\]}
\noindent where $\epsilon_t$ is a stationary process. When $\mu = 0 $ the process is called pure random walk and when $\mu \neq 0$  the process is called random walk with drift.
\end{block}
%\begin{block}{Deterministic trend}
% Deterministic trend or trend stationary
%processes have the following form:
%{\color{blue}
%\[
%y_t = f(t) + \epsilon_{t}
%\]}
%\noindent where $\epsilon_{t}$ represents a stationary error term (with mean 0 and variance $\sigma^2$) and $f(t)$ is a deterministic function of time.
%\end{block}
\end{frame}

%\begin{frame}[t,fragile]{Random walk Illustration}
% \begin{itemize}
%  \item<1-> The drunkard's walk
%  \item<2-> The dog's walk
%  %\item<3-> The drunk and dog's walk
% \end{itemize}
% 
%\begin{columns}
%\begin{column}{.4\linewidth}<1->
%\begin{center}
%        \begin{tikzpicture}[overlay]
%            \node[yshift=-50] (img2) {\includegraphics[scale=1.4]{img/drunk}};
%        \end{tikzpicture}
%\end{center}
%\end{column}
%\begin{column}{.08\linewidth}<2->
%        \begin{tikzpicture}[overlay]
%            \node[yshift=-50] (img2) {\includegraphics[scale=0.1]{img/dog}};
%        \end{tikzpicture}
%\end{column}
%%\begin{column}{.5\linewidth}<3->
%%    \begin{center}
%%        \begin{tikzpicture}[overlay]
%%            \node[yshift=-50] (img2) {\includegraphics[scale=0.1]{img/mananddog}};
%%        \end{tikzpicture}
%%    \end{center}
%%\end{column}
%\end{columns}
%\end{frame}
%
%\begin{frame}[t,fragile]{Random walk and Cointegration Illustration}
% \begin{itemize}
%  \item The drunk and dog's walk
% \end{itemize}
%    \begin{center}
%        \begin{tikzpicture}[overlay]
%            \node[yshift=-60] (img2) {\includegraphics[scale=0.7]{img/mananddog}};
%        \end{tikzpicture}
%    \end{center}
%\end{frame}

\begin{frame}
\frametitle{Irregularly spaced over time}
\begin{figure}
\includegraphics[width=\paperwidth]{img/hft-ticks-zoom.png}
\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{Motivation}

The relationship between non-stationary time series can be modelled if a stationary linear combination is shown to exist.
%Financial time series are mainly non-stationary.  However, one or more linear
%combinations of non-stationary time series can be stationary even though
%individually they are not. 
When this happens it is said they have a long-run
equilibrium relationship and the variables are {\color{red} cointegrated}.
\end{frame}

\begin{frame}
\frametitle{Cointegration}
\begin{block}{Cointegration definition}
Let {\color{blue}$\mathbf{y} = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$} be a set of $l$
time series of order I(1) which are said to be cointegrated if a vector
$\beta=[\beta(1),\dots,\beta(l)]^\top \in \mathbb{R}^l$  exists such that the
time series,

\begin{equation*}
 \mathbf{Z}_t:= \beta^\top \mathbf{y} = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
 \text{I(0)}\, .
\end{equation*}
\end{block}
In other words, a set of I(1) variables is said to be cointegrated if
a linear combination of them exists which is I(0).

@@

The Vector Error correction model (VECM) is used to model cointegrated time series.

{\Large
\begin{equation*}
 \Delta \mathbf{y}_t = \underbrace{
        \tikz[baseline]{
            \node[fill=red!20,ellipse,anchor=base] (t1)
            {$ \Omega\mathbf{y}_{t-1}$};
        }}_\text{Error correction term} +
        \underbrace{
        \tikz[baseline]{
            \node[fill=blue!20,anchor=base] (t2)
            {$\sum_{i=1}^{p-1} \phi_i^* \Delta \mathbf{y}_{t-i} $};
        }}_\text{Autorregresive term} +
        \underbrace{
        \tikz[baseline]{
            \node[fill=green!20,anchor=base] (t3)
            {$c$};
        }}_\text{Intercept}
        +
        \underbrace{
        \tikz[baseline]{
            \node[fill=yellow!20,anchor=base] (t3)
            {$\mathbf{\epsilon}_t $};
        }}_\text{i.i.d}
\end{equation*}
}
The columns of $\beta$ contains the cointegration vectors and the rows of
$\alpha$ correspond with the adjusted vectors. $\beta$ is obtained by Johansen
procedure whereas $\alpha$ has to be determined as a
variable in the VECM.

\begin{frame}
\frametitle{VECM matrix form}
\small
\begin{equation*}
\underbrace{
      \begin{bmatrix}
       \mathbf{\Delta y}_{p+1}  \\ 
       \mathbf{\Delta y}_{p+2}  \\ 
       \vdots                   \\ 
       \mathbf{\Delta y}_N      
      \end{bmatrix}}_{\mathbf{B} } =
\arraycolsep=1.4pt  
\underbrace{\begin{bmatrix}
   \alpha^\top & \phi^*_1 & \dots & \phi^*_{p-1} &\mathbf{c}   
  \end{bmatrix}}_{\mathbf{X}^\top}
\underbrace{\begin{bmatrix}
 \beta\mathbf{y}_p      & \beta\mathbf{y}_{p+1}   & \dots&   \beta\mathbf{y}_{N-1}   \\
 \mathbf{\Delta y}_{p}   & \mathbf{\Delta y}_{p+1}& \dots &  \mathbf{\Delta y}_{N-1} \\
 \mathbf{\Delta y}_{p-1} & \mathbf{\Delta y}_{p}  & \dots &  \mathbf{\Delta y}_{N-2}   \\
 \vdots                  & \vdots                 & \ddots&  \vdots                   \\
 \mathbf{\Delta y}_{2}   & \mathbf{\Delta y}_{3} & \dots &   \mathbf{\Delta y}_{N-p+1} \\
 1                      & 1                       & \dots     & 1   
 \end{bmatrix}}_{\mathbf{A}^\top}
+
\underbrace{\begin{bmatrix}
              \mathbf{\epsilon}_{p+1} \\ 
              \mathbf{\epsilon}_{p+2} \\ 
              \vdots \\ 
              \mathbf{\epsilon}_N
             \end{bmatrix}}_{\mathbf{E}^\top}
\end{equation*}
\end{frame}

\begin{frame}
\frametitle{$\Omega$ matrix properties}
The matrix $\Omega$ has the following properties:
\begin{itemize}
\item If $\Omega = 0$ there is no cointegration
\item If $rank(\Omega)=l$ i.e full rank, then the time series are not
I(1) but stationary
\item If $rank(\Omega)=r,\quad 0 < r < l$ then, there is cointegration
and the matrix $\Omega$ can be expressed as $\Omega =
\alpha \beta^\top$, where $\alpha$ and $\beta$ are $(l \times r)$
matrices and $rank(\alpha)=rank(\beta)=r$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{VECM limitations}
The Vector Error correction model (VECM) is used to model cointegrated time series but only with batch data and rarely used with high frequency data  mainly due to computational limitations:
\begin{itemize}
\item Calculation of the cointegration vectors
\item VECM parameters are obtained using the ordinary least squares (OLS) method.
\end{itemize}
\end{frame}

\section{Proposal}
\begin{frame}
\frametitle{Proposal}
\begin{block}{Proposal}
To adapt VECM to be used with high frequency data, where the best parameters are updated when new data arrives.
\end{block}
\begin{block}{Hypothesis}
An algorithm based on cointegration and high performance computing will allow faster forecasting
algorithms for financial time series to be obtained while maintaining good accuracy levels.
\end{block}
\end{frame}

\section{First Contribution}
\begin{frame}
\frametitle{First approach}
An adaptive Vector Error Correction algorithm that:
\begin{itemize}
\item considers the last most recent data (windows size L)
\item finds optimal parameters of VECM in that window.
\item the optimisation is made in a distributed environment.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{First approach idea}
\begin{figure}
\includegraphics[width=0.65\textwidth]{img/slidingwindow}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{How to choose L?}
  \begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
   \includegraphics[width=0.65\textwidth]{img/51_Fig1}
  \caption[Distribution of the number of cointegration vectors using $p=1$ lags]{Distribution of the number of cointegration vectors using $p=1$ lags.
  Four possible values for windows size $L$ are shown (2, 6, 10 and 14) hours (1
  hour = 360 data points).}
  \label{fig:hists}
\end{figure}
\end{frame}

\subsection{Data}
\begin{frame}
\frametitle{Data}
\begin{itemize}
\item Tests were carried out using four foreign exchange rates all related to
USD: EURUSD, GBPUSD, USDCHF and USDJPY. 
\item This data was collected from Dukascopy, a free
database which gives access to the Swiss Foreign Exchange marketplace.
\item The tests were done using 10-seconds frequency from ask prices
from the 11th to the 15th of August 2014. 
%\item Since one day corresponds to 8640 data points and we used 5 days of data, we have 43,200 data points in total.
\end{itemize}
 \begin{figure}[!h]
    \includegraphics[width=0.2\textwidth]{img/dukascopy}
 \end{figure}
\end{frame}

@@

%\begin{frame}
%\frametitle{PC vs MSE}
%\begin{figure}[ht!]
%  %\vspace{-0.8cm}
%  \centering
%  \includegraphics[width=0.7\textwidth]{img/MSE-offset20520-p-2-freq-10s}
%  \caption{MSE versus the percentage of cointegration considering 1000
%  iterations. }
%  \label{fig:cointvsmse}
%\end{figure}
%\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{AVECM}
\begin{frame}
\frametitle{Adaptive VECM algorithm (AVECM)}
\begin{itemize}
\item We proposed to choose $L$ and $p$ in order to maximise the percentage of
cointegration $PC$ in the near past. This process was done every time that new data was processed. 
\item To use a distributed environment to make this parameters search, since this is the most expensive routine.
\end{itemize}
\end{frame}

@@

\section{Conclusions}
\subsection{Uno}
\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
\item Cointegration relations change with time. We empirically showed that the Johansen method is sensitive to the number of lags but also to the amount of data considered.
\item We found that out-of-sample forecast performance MSE is related to the {\em percentage of cointegration\/}.  
\item VECM improves performance measures by finding
parameters of $L$ and $p$ maximising the percentage of cointegration.
\item Increasing the number of parameters will always lead to
better performance measures.
\item The parallel implementation allowed the execution times to be reduced
more than 9 times and therefore a response time was obtain before 10
seconds. 
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Integration }
\begin{block}{Integration of order $d$}
A time series $\mathbf{y}$ is said to be integrated of order $d$ if after
differentiating the variable $d$ times, we get an stationary process, more precisely:
\[
(1-L)^d \mathbf{y} \sim \text{I(0)} \, ,
\]
\noindent where I(0) is a stationary time series and $L$ is the lag operator:
\[
(1-L)\mathbf{y} = \Delta \mathbf{y}=\mathbf{y}_t  -\mathbf{y}_{t-1} \quad \forall t
\]
\end{block}
\end{frame}

>>>>

\documentclass{beamer}
%\usepackage[spanish,activeacute]{babel}
\usetheme{metropolis}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{color}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{array}
%\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{booktabs}
\usetikzlibrary{shapes,arrows}
%\usepackage{multicol}
\usepackage{hhline}
\usepackage{pgf}
\usepackage{algorithm} 
\usepackage{algorithmic}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\def\R{{\mathbb R }}
\def\N{{\mathbb N }}

\definecolor{BlueTOL}{HTML}{222255}
\definecolor{BrownTOL}{HTML}{f43d00}%666633}
\definecolor{GreenTOL}{HTML}{225522}
\setbeamercolor{normal text}{fg=BlueTOL,bg=white}
\setbeamercolor{alerted text}{fg=BrownTOL}
\setbeamercolor{example text}{fg=GreenTOL}

@@

\setbeamercolor{block title alerted}{use=alerted text,
    fg=alerted text.fg,
    bg=alerted text.bg!80!alerted text.fg}
\setbeamercolor{block body alerted}{use={block title alerted, alerted text},
    fg=normal text.fg,
    bg=block title alerted.bg!50!alerted text.bg}
\setbeamercolor{block title example}{use=example text,
    fg=example text.fg,
    bg=example text.bg!80!example text.fg}
\setbeamercolor{block body example}{use={block title example, example text},
    fg=example text.fg,
    bg=block title example.bg!50!example text.bg}

@@

\author[Paola Arce]{Paola Arce \\ Thesis advisor: Luis Salinas}
\date{December 11, 2017}
\title[Thesis defense]{Fast and adaptive cointegration based model for forecasting financial time series}
\subtitle{Seminario Inform\'atica CCTVal}

@@

\begin{frame}[plain]
\titlepage
\begin{tikzpicture}[overlay, remember picture]
\node[above right=0.4cm and 1.2cm of current page.south west] {\includegraphics[width=2.5cm]{img/UTFSM}};
%\node[above =0.5cm of current page.south] {\includegraphics[width=2cm]{img/DIT}};
\node[above left=0.5cm and .8cm of current page.south east] {\includegraphics[width=5cm]{img/logo-di}};
\end{tikzpicture}

\end{frame}
%\usebackgroundtemplate{ }

\begin{frame}
\frametitle{Outline}
\begin{itemize}
\item Motivation: High Frequency trading
\item Cointegration and the VECM model
\item Challenges and Hypothesis
\item Contributions
\item Experiments
\item Conclusions
\item Future work
\end{itemize}
\end{frame}

\section{Motivation}

@@

\begin{frame}
\frametitle{High Frequency data}
High frequency time series are commonly:
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item Non-stationary
\item Irregularly spaced over time.
%\item Heterogeneity
\item Heteroscedastic
\item Potentially cointegrated
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=0.35\paperwidth]{img/forecast}
\end{figure}
\end{column}
\end{columns}
\vspace{5mm}
This imply that classical statistical models can't be used. Therefore, financial time series forecasting is still a challenge.
\end{frame}

@@

\begin{frame}
\frametitle{Stationarity: consequences}
One of the consequences of stationary processes is how they recover from a shock. A shock represents an unexpected change in a variable in a particular period of time. For stationary time series, shocks to the system will gradually die away. 
\begin{figure}
\includegraphics[width=0.6\paperwidth]{img/shock}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Non-stationary processes}
There are different types of non-stationary time series models often found in economics. Unit root process is one of them commonly used to model prices:
\begin{block}{Unit root process}
{\color{red}Unit root or I(1)} processes are also called stochastic trend if they have the following form:
{\color{blue}
\[
y_t = \mu + y_{t-1} + \epsilon_t
\]}
\noindent where $\epsilon_t$ is a stationary process. When $\mu = 0 $ the process is called {\bf pure random walk} and when $\mu \neq 0$  the process is called {\bf random walk with drift}.
\end{block}
\end{frame}

\begin{frame}
\frametitle{Integration }
\begin{block}{Integration of order $d$}
A time series $\mathbf{y}$ is said to be I(d), or integrated of order $d$, if after
differentiating the variable $d$ times, we get an stationary process, more precisely:
\[
{\color{blue}
(1-L)^d \mathbf{y} \sim \text{I(0)}} \, ,
\]
\noindent where I(0) is a stationary time series and $L$ is the lag operator:
\[
(1-L)\mathbf{y} = \Delta \mathbf{y}=\mathbf{y}_t  -\mathbf{y}_{t-1} \quad \forall t
\]
\end{block}
\end{frame}

\begin{frame}
\frametitle{Irregularly spaced over time}
\begin{figure}
\includegraphics[width=0.8\paperwidth]{img/hft-ticks-zoom.png}
\end{figure}
\end{frame}

@@

\section{Cointegration and the VECM model}
\begin{frame}
\frametitle{Cointegration}
\begin{figure}
\includegraphics[width=0.8\paperwidth]{img/motivation}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Cointegration definition}

\begin{block}{Cointegration}
The relationship between non-stationary time series can be modelled if a {\bf stationary linear combination} is shown to exist.
When this happens it is said they have a {\bf long-run
equilibrium} relationship and the variables are {\color{red} cointegrated}.
\end{block}

\begin{frame}
\frametitle{Cointegration: formal definition}
\begin{block}{Cointegration definition}
Let {\color{blue}$\mathbf{y} = \{\mathbf{y}^1, \dots, \mathbf{y}^l\}$} be a set of $l$
time series of order I(1) which are said to be cointegrated if a vector
${\color{blue}\boldsymbol\beta=[\beta(1),\dots,\beta(l)]^\top} \in \mathbb{R}^l$  exists such that the
time series,

\begin{equation*}
 \mathbf{Z}_t:= \boldsymbol \beta^\top \mathbf{y} = \beta(1) \mathbf{y}^1 + \dots + \beta(l) \mathbf{y}^l \sim
 \text{I(0)}\, .
\end{equation*}
\end{block}
%In other words, a set of I(1) variables is said to be cointegrated if
%a linear combination of them exists which is I(0).

@@

The Vector Error correction model (VECM) is used to model cointegrated time series:

{\Large
\begin{equation*}
 \Delta \mathbf{y}_t = \underbrace{
        \tikz[baseline]{
            \node[fill=red!20,ellipse,anchor=base] (t1)
            {$ \Omega\mathbf{y}_{t-1}$};
        }}_\text{Error correction term} +
        \underbrace{
        \tikz[baseline]{
            \node[fill=blue!20,anchor=base] (t2)
            {$\sum_{i=1}^{p-1} \phi_i^* \Delta \mathbf{y}_{t-i} $};
        }}_\text{Autorregresive term} +
        \underbrace{
        \tikz[baseline]{
            \node[fill=green!20,anchor=base] (t3)
            {$c$};
        }}_\text{Intercept}
        +
        \underbrace{
        \tikz[baseline]{
            \node[fill=yellow!20,anchor=base] (t3)
            {$\mathbf{\epsilon}_t $};
        }}_\text{i.i.d}
\end{equation*}
}
\noindent where $p$ is the number of lags and $\boldsymbol \Omega = \boldsymbol \alpha \boldsymbol \beta^\top$. The columns of $\boldsymbol\beta$ contains the cointegration vectors and the rows of $\boldsymbol\alpha$ correspond with the adjusted vectors.
\end{frame}

\begin{frame}
\frametitle{$\boldsymbol \Omega$ matrix properties}
The matrix $\boldsymbol \Omega$ has the following properties:
\begin{itemize}
\item If ${\color{blue}\boldsymbol \Omega = 0}$ there is no cointegration
\item If ${\color{blue}rank(\boldsymbol\Omega)=l}$ i.e full rank, then the time series are not
I(1) but stationary
\item If ${\color{blue}rank(\boldsymbol\Omega)=r,\quad 0 < r < l}$ then, there is cointegration.
%where $\boldsymbol\alpha$ and $\boldsymbol\beta$ are $(l \times r)$
%matrices and $rank(\boldsymbol\alpha)=rank(\boldsymbol\beta)=r$.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{VECM limitations}

\begin{alertblock}{VECM limitations}
The Vector Error correction model is used to model cointegrated time series but only with {\bf batch data and rarely used with high frequency data} mainly due to {\bf computational limitations}:
\begin{itemize}
\item Calculation of the cointegration vectors $\boldsymbol \beta$ is obtained by the Johansen procedure which is of order $O(n^3)$.
%whereas $\alpha$ has to be determined as a variable in the VECM.
\item VECM parameters are obtained using the ordinary least squares ({\bf OLS}) method.
\end{itemize}
\end{alertblock}
\end{frame}

\section{Challenges and Hypothesis}
\begin{frame}
\frametitle{Challenge}
\begin{block}{Proposal}
To adapt VECM to be used with high frequency data, where the best parameters are updated when new data arrives.
\end{block}
\begin{exampleblock}{Hypothesis}
An algorithm based on cointegration and high performance computing will allow faster forecasting
algorithms for financial time series to be obtained while maintaining good accuracy levels.
\end{exampleblock}
\end{frame}

\section{Contributions}
\begin{frame}
\frametitle{First contribution: Adaptive VECM}
\begin{figure}
\includegraphics[width=\textwidth]{img/slidingwindow}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{First approach idea}
An adaptive Vector Error Correction (AVECM) algorithm that:
\begin{itemize}
\item considers the last most recent data (windows size L)
\item finds optimal parameters of VECM in that window.
\item the optimisation is made in a distributed environment.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{How to choose L?}
  \begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
   \includegraphics[width=0.85\textwidth]{img/51_Fig1}
  \caption[Distribution of the number of cointegration vectors using $p=1$ lags]{Distribution of the number of cointegration vectors using $p=1$ lags.}
  %Four possible values for windows size $L$ are shown (2, 6, 10 and 14) hours (1 hour = 360 data points).}
  \label{fig:hists}
\end{figure}
\end{frame}

@@

\begin{frame}
\frametitle{Data}
\begin{columns}
\begin{column}{0.7\textwidth}
\begin{itemize}
\item This data was collected from {\bf Dukascopy}, a free
database which gives access to the {\bf Swiss Foreign Exchange marketplace}.
\item Tests were carried out using four foreign exchange rates all related to
USD: {\bf EURUSD, GBPUSD, USDCHF} and {\bf USDJPY}. 
\item The tests were done using 10-seconds frequency from ask prices
from the 11th to the 15th of August 2014. 
%\item Since one day corresponds to 8640 data points and we used 5 days of data, we have 43,200 data points in total.
\end{itemize}
\end{column}
\begin{column}{0.3\textwidth}
 \begin{figure}[!h]
    \includegraphics[width=\textwidth]{img/dukascopy}
 \end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{PC vs MSE}
\begin{figure}[ht!]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.7\textwidth]{img/51_Fig2}
  \caption{MSE versus the percentage of cointegration considering 1000
  iterations. }
  \label{fig:cointvsmse}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{frame}
\frametitle{Adaptive VECM algorithm (AVECM)}
\begin{itemize}
\item We proposed to choose $L$ and $p$ in order to maximise the percentage of
cointegration $PC$ in the near past. This process was done every time that new data was processed. 
\item To use a distributed environment to make this parameters search, since this is the most expensive routine.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Conclusions}
\begin{itemize}
\item Cointegration relations change with time. We empirically showed that the Johansen method is sensitive to the number of lags but also to the amount of data considered.
\item We found that out-of-sample forecast performance MSE is related to the {\em percentage of cointegration\/}.  
\item VECM improves performance measures by finding
parameters of $L$ and $p$ maximising the percentage of cointegration.
\item Increasing the number of parameters will always lead to
better performance measures.
\item The parallel implementation allowed the execution times to be reduced
more than 9 times and therefore a response time was obtain before 10
seconds. 
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{VECM(p) matrix form}
If we have $N$ data points, with $N>p$, we can estimate the model parameters $\mathbf{X}$ using the following matrix form:
\small
\begin{equation*}
\underbrace{
      \begin{bmatrix}
       \mathbf{\Delta y}_{p+1}  \\ 
       \mathbf{\Delta y}_{p+2}  \\ 
       \vdots                   \\ 
       \mathbf{\Delta y}_N      
      \end{bmatrix}}_{\mathbf{B} } =
\arraycolsep=1.4pt  
\underbrace{\begin{bmatrix}
   \alpha^\top & \phi^*_1 & \dots & \phi^*_{p-1} &\mathbf{c}   
  \end{bmatrix}}_{\mathbf{X}^\top}
\underbrace{\begin{bmatrix}
 \beta\mathbf{y}_p      & \beta\mathbf{y}_{p+1}   & \dots&   \beta\mathbf{y}_{N-1}   \\
 \mathbf{\Delta y}_{p}   & \mathbf{\Delta y}_{p+1}& \dots &  \mathbf{\Delta y}_{N-1} \\
 \mathbf{\Delta y}_{p-1} & \mathbf{\Delta y}_{p}  & \dots &  \mathbf{\Delta y}_{N-2}   \\
 \vdots                  & \vdots                 & \ddots&  \vdots                   \\
 \mathbf{\Delta y}_{2}   & \mathbf{\Delta y}_{3} & \dots &   \mathbf{\Delta y}_{N-p+1} \\
 1                      & 1                       & \dots     & 1   
 \end{bmatrix}}_{\mathbf{A}^\top}
+
\underbrace{\begin{bmatrix}
              \mathbf{\epsilon}_{p+1} \\ 
              \mathbf{\epsilon}_{p+2} \\ 
              \vdots \\ 
              \mathbf{\epsilon}_N
             \end{bmatrix}}_{\mathbf{E}^\top}
\end{equation*}
\end{frame}
